The promise of community-driven preprints in ecology and evolution

### 2024-05-03 ms/ms.qmd
More corrections

```{r, fig-summary}
#| label: fig-summary
#| fig-cap: Summary of articles posted to *EcoEvoRxiv*. A) Number of articles (preprints and postprints) published monthly to *EcoEvoRxiv* between 2018 and 2023. *EcoEvoRxiv* was established in June 2018 before the soft launch in November 2018. Notable milestones include when *EcoEvoRxiv* transitioned to the California Digital Library (CDL), the acceptance of Spanish and Portuguese preprints and postprints and when the first IUCN Red list Ecosystem report was accepted; B) Countries that are posting articles (preprints and postprints) to *EcoEvoRxiv* as inferred from the country of affiliation of the submitting author; C) Taxa used / covered in the articles posted to *EcoEvoRxiv* (*n* = 1080); D) Types of preprints accepted on *EcoEvoRxiv* (*n* = 620). E) Career stage of authors posting to *EcoEvoRxiv* along with the gender of the submitting author.
#| echo: false
#| warning: false
#| message: false

@@

After removing `r dim(dupl)[1]` duplicate titles – suggesting a few authors created multiple submissions for the same preprint rather than updating an existing submission – we had data for a total of `r pr` articles with ~55--60 preprints published each month in the last two years ([@fig-summary]A).  *EcoEvoRxiv* hosts papers from authors based in `r length(unique(countries$country))-1` countries, with 90% of the articles coming from just `r nrow(most_use)-1` countries. North America, Australia, and Europe upload the most preprints with much fewer coming from countries in Africa, Central America and parts of Asia ([@fig-summary]B).  Articles covered all major taxonomic groups, with the most common being vertebrates (`r sum_taxa$perc[6]`%), plants (`r sum_taxa$perc[1]`%) and invertebrates (`r sum_taxa$perc[5]`%) ([@fig-summary]C).

### *Diversifying articles types on **EcoEvoRxiv**: overcoming the 'grey literature' problem*

@@

Accepting a greater diversity of article types allows *EcoEvoRxiv* to begin dealing with the 'grey literature' problem whereby data that are relevant for research syntheses are not published in typical peer-reviewed journals[@haddaway2015shades; @haddaway2020eight]. This is reflected by `r nt`% of articles on *EcoEvoRxiv* that could be considered ‘grey literature’ in ecology and evolutionary biology: books, book chapters, reports, and other research output types. Currently, *EcoEvoRxiv* does not host many reports (`r sum_types[sum_types$type == "report", "perc"]`%), particularly from government or industry. We would encourage authors to consider posting these types of articles to ensure they are openly accessible and more easily found. *EcoEvoRxiv* articles are given a unique DOI and indexed on Google Scholar. *EcoEvoRxiv* has already formed fruitful partnerships with the International Union for Conservation of Nature (IUCN) as a place where IUCN reports can be hosted to make them more easily accessible to the broader research community. For example, IUCN Red Ecosystem Reports are now posted to *EcoEvoRxiv* and our community has been able to work closely with the IUCN to ensure these documents meet the IUCN requirements.

@@

*EcoEvoRxiv* is the only preprint server to date breaking down language barriers to scientific communication; currently accepting English, Spanish, Portuguese and French manuscripts. Such initiatives are incredibly important if we are to begin filling global voids of scientific knowledge. However, multilingual initiatives have been slow to take off on *EcoEvoRxiv*, with only a few Spanish preprints posted to date, and none in other accepted languages (e.g., French and Portuguese). Part of the challenge in getting authors to submit non-English preprints is the lack of awareness of *EcoEvoRxiv* in non-English speaking countries and the fact that it accepts preprints in different languages, cultural differences in the perception of preprints, and a strong reliance on traditional publishing models which typically mandate publishing in English.

@@

### *Science takes time, but publication could take longer*
Increased competition in science has raised the bar with respect to the amount of data required for publication [@vale2015accelerating]. This is a good outcome if it results in higher impact research that better clarifies our understanding of the natural world, but it does come at a cost for the rate at which research findings can be disseminated [@vale2015accelerating]. Preprinting has been proposed as a way to disseminate research more quickly as it can take a long time before results are ultimately published [@bourne2017ten; @vale2015accelerating]. However, data on the time to publication is sparse, and the real benefit of preprints in this context would be reinforced if preprinting can dramatically speed up dissemination. We evaluated how long it takes to publish a research paper in ecology and evolution by ascertaining the time between when a preprint was first posted and finally accepted in a research journal. In total, `r unpub` papers were still unpublished out of the `r nrow(data2)` (`r (unpub / nrow(data2))*100`%) when these data were collected. Astonishingly, the maximum time between preprint submission and publication was `r sum_preprints$max` days or `r sum_preprints$max/365` years ([@fig-pubsummary]A), and the median time to publication was `r sum_preprints$median` days (mean = `r sum_preprints$mean`; SD = `r sum_preprints$sd` days). Our results largely confirm the long timespans most authors are dealing with to get their research published after the paper is written.

>>>>

```{r, fig-summary}
#| label: fig-summary
#| fig-cap: Summary of articles posted to *EcoEvoRxiv*. A) Number of articles (preprints and postprints) published monthly to *EcoEvoRxiv* between 2018 and 2023. *EcoEvoRxiv* was established in June 2018 before the soft launch in November 2018. Notable milestones include when *EcoEvoRxiv* transitioned to the California Digital Library (CDL), the acceptance of Spanish and Portuguese preprints and postprints and when the first IUCN Red list Ecosystem report was accepted; B) Countries that are posting articles (preprints and postprints) to *EcoEvoRxiv* as inferred from the country of affiliation of the submitting author; C) Taxa used / covered in the articles posted to *EcoEvoRxiv* (*n* = 1080 papers); D) Types of preprints accepted on *EcoEvoRxiv* (*n* = 620 papers). E) Career stage of authors posting to *EcoEvoRxiv* along with the gender of the submitting author.
#| echo: false
#| warning: false
#| message: false

@@

After removing five duplicate titles – suggesting a few authors created multiple submissions for the same preprint rather than updating an existing submission – we had data for a total of `r pr` articles with ~55--60 preprints published each month in the last two years ([@fig-summary]A).  *EcoEvoRxiv* hosts papers from authors based in `r length(unique(countries$country))-1` countries, with 90% of the articles coming from just `r nrow(most_use)-1` countries. North America, Australia, and European countries upload the most preprints with much fewer coming from countries in Africa, Central America, and parts of Asia ([@fig-summary]B).  Articles covered all major taxonomic groups, with the most common being vertebrates (`r sum_taxa$perc[6]`%), plants (`r sum_taxa$perc[1]`%), and invertebrates (`r sum_taxa$perc[5]`%) ([@fig-summary]C).

### *Diversifying article types on **EcoEvoRxiv**: overcoming the 'grey literature' problem*

@@

Accepting a greater diversity of article types allows *EcoEvoRxiv* to begin dealing with the 'grey literature' problem whereby data that are relevant for research syntheses are not published in typical peer-reviewed journals[@haddaway2015shades; @haddaway2020eight]. This is reflected by `r nt`% of articles on *EcoEvoRxiv* that could be considered ‘grey literature’ in ecology and evolutionary biology: books, book chapters, reports, and other research output types. Currently, *EcoEvoRxiv* does not host many reports (`r sum_types[sum_types$type == "report", "perc"]`%), particularly from governments or industry. We would encourage authors to consider posting books, book chapters and reports to ensure they are openly accessible and more easily found. *EcoEvoRxiv* articles are given a unique DOI and indexed on Google Scholar. *EcoEvoRxiv* has already formed fruitful partnerships with the International Union for Conservation of Nature (IUCN) as a place where IUCN reports can be hosted to make them more easily accessible to the broader research community. For example, IUCN Red Ecosystem Reports are now posted to *EcoEvoRxiv* and our community has been able to work closely with the IUCN to ensure these documents meet the IUCN requirements.

@@

*EcoEvoRxiv* is the only preprint server to date breaking down language barriers to scientific communication; currently accepting English, Spanish, Portuguese and French manuscripts. Such initiatives are incredibly important if we are to begin filling global voids of scientific knowledge. However, multilingual initiatives have been slow to take off on *EcoEvoRxiv*, with only a few Spanish preprints and a single Portuguese preprint posted to date. Part of the challenge in getting authors to submit non-English preprints is the lack of awareness of *EcoEvoRxiv* in non-English speaking countries and the fact that it accepts preprints in different languages, cultural differences in the perception of preprints, and a strong reliance on traditional publishing models which typically mandate publishing in English.

@@

### *Science takes time, but publication could take longer*
Increased competition in science has raised the bar with respect to the amount of data required for publication [@vale2015accelerating]. This is a good outcome if it results in more comprehensive research that better clarifies our understanding of the natural world, but it does come at a cost for the rate at which research findings can be disseminated [@vale2015accelerating]. Preprints have been proposed as a way to disseminate research more quickly as it can take a long time before results are ultimately published [@bourne2017ten; @vale2015accelerating]. However, data on the time to publication is sparse, and the real benefit of preprints in this context would be reinforced if preprinting can dramatically speed up dissemination. We evaluated how long it takes to publish a research paper in ecology and evolution by ascertaining the time between when a preprint was first posted and finally accepted in a peer-reviewed journal. In total, `r unpub` papers were still unpublished out of the `r nrow(data2)` (`r (unpub / nrow(data2))*100`%) when these data were collected. Not all of these papers, however, are anticipated to be published in a peer-reviewed journal (e.g., reports). The maximum time between preprint submission and publication was `r sum_preprints$max` days or `r sum_preprints$max/365` years ([@fig-pubsummary]A), and the median time to publication was `r sum_preprints$median` days (mean = `r sum_preprints$mean`; SD = `r sum_preprints$sd` days). Our results largely confirm the extended timeframes that most authors experience between writing their research papers and their publication.
